Compiler and Runtime Support for Functions-as-a-Service (FaaS)

### 2021-12-09 paper/paper.tex
[paper]: add figures for static analysis and pool transformation, start allocator subsection, fix content

\subsection{Profiler}
The profiler has the goal of understanding information about dynamic memory \ALLS\ and 
\DALLS\ during execution (for instance, the number of \ALLS\, \DWS\, common \ALL 
sizes, etc.) as well as standard information about the "hotness" of functions and
loops at runtime. Using this data, the profiler can inform the compiler during 
instrumentation in the second pass over the application (see \ref{ca}). It can also 
inform the user with a better estimate of memory usage for a future FaaS instance or 
function invocation.

@@

\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/pins.png} 
        \caption{Example compiler instrumentation for profiler callbacks in the 
        context of simplified LLVM IR and \textit{libc} \texttt{malloc}
        and \texttt{free}}  
	    \label{fig:pins}
    \end{minipage}\hfill
\end{figure}

@@

\subsection{Static Analyses}
The system also supports static analyses of \ALLS\ and \DALLS\ in the IR for workload $W$. 
In particular, the compiler analyzes 1) \ALL\ sizes for all \ALLS\, 2) loops to unroll 
and functions to inline so that 1) is more feasible to analyze, and 3) uses of \ALL\ and 
\DALLS . All analyses informs the compiler about analyzable static allocations.

@@

An "easy-to-deduce" or \textit{statically analyzable} \SIZE\ is one that is 
simply a constant integer. All other \SIZES\ are \textit{analyzable with
assistance}, where the size can apparently be deduced during runtime. These 
categories are mutually exclusive. The compiler records all statically analyzable 
\SIZES .

Some \SIZES\ that are analyzable with assistance are worth recording, however. In
particular, if an \SIZE\ is dependent on function arguments, the function (and
its parent functions) can be continually inlined until the size is statically
analyzable. Consequently, the compiler marks functions that are candidates for
aggressive inlining (compared to a standard inliner in \texttt{-O3}). At an
extreme, aggressive inlining can reach a point where it may be beneficial to
inline the entry-point function of $W$. This is especially common in small 
FaaS workloads where \ALLS\ towards the beginning of the workload havew \SIZES\
dependent on the inputs. Obviously, inlining the entry-point function is not 
possible, but we can borrow concepts from JIT compilation or binary translation
to possibly embed the inputs into the binary itself. This process can occur
right before the $W$ is invoked in the FaaS instance. A subset of this system's compiler
can analyze and transform $W$ as necessary. This solution is theoretical and 
not applied in this work.

Additionally, \SIZES\ that exists in a loop and are dependent on loop-invariant
values are also recorded. Here, the loop itself can be aggressively unrolled and
the \ALLS\ can be merged or replaced with one of the methods from the custom 
allocator. It is also important that the size is dependent on a loop-invariant 
value, or else the compiler cannot affirmatively analyze multiples of the size or
determine an appropriate unroll factor (unrolling too much without knowledge of
the \SIZE\ may be sub-optimal).

@@

\subsection{Custom Allocator} \label{ca}

>>>>

\subsection{Profiler}
The profiler has the goal of understanding information about dynamic memory \ALLS\ and 
\DALLS\ during execution (for instance, the number of \ALLS\, \DWS\, common \ALL 
sizes, etc.) as well as typical information about critical paths or the "hotness" of functions and
loops at runtime that is often use to aid optimization [?]. Using this data, the profiler can inform the compiler during 
instrumentation in the second pass over the application (see \ref{ca}). It can also 
inform the user with a better estimate of memory usage for a future FaaS instance or 
function invocation.

@@

\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/pins.png} 
        \caption{Example compiler instrumentation for profiler callbacks in the 
        context of simplified LLVM IR and \textit{libc} \texttt{malloc}
        and \texttt{free}}  
	    \label{fig:pins}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.77\textwidth]{figs/sizes.png} 
        \caption{Example allocations that the static analysis can target in the 
        context of simplified LLVM IR and \textit{libc} \texttt{malloc}}  
	    \label{fig:sizes}
    \end{minipage}\hfill
\end{figure}

@@

\subsection{Static Analyses} \label{sa}
The system also supports static analyses of \ALLS\ and \DALLS\ in the IR for workload $W$. 
In particular, the compiler analyzes 1) \ALL\ sizes for all \ALLS\, 2) loops to unroll 
and functions to inline so that 1) is more feasible to analyze, and 3) uses of \ALL\ and 
\DALLS . All analyses informs the compiler about analyzable static allocations.

@@

An "easy-to-deduce" or \textit{statically analyzable} \SIZE\ is one that is 
simply a constant integer. All other \SIZES\ are \textit{analyzable with
assistance}, where the size can apparently be deduced during runtime. These 
categories are mutually exclusive. The compiler records all statically analyzable 
\SIZES . An example is shown in Line 2 of Figure~\ref{fig:sizes}.

Some \SIZES\ that are analyzable with assistance are worth recording, however. In
particular, if an \SIZE\ is dependent on function arguments, the function (and
its parent functions) can be continually inlined until the size is statically
analyzable. Consequently, the compiler marks functions that are candidates for
aggressive inlining (compared to a standard inliner in \texttt{-O3}). An example of
this is shown in Line 3 of Figure ~\ref{fig:sizes}. At an
extreme, aggressive inlining can reach a point where it may be beneficial to
inline the entry-point function of $W$. This is especially common in small 
FaaS workloads where \ALLS\ towards the beginning of the workload havew \SIZES\
dependent on the inputs. Obviously, inlining the entry-point function is not 
possible, but we can borrow concepts from JIT compilation or binary translation
to possibly embed the inputs into the binary itself. This process can occur
right before the $W$ is invoked in the FaaS instance. A subset of this system's compiler
can analyze and transform $W$ as necessary. This solution is theoretical and 
not applied in this work.

Additionally, \SIZES\ that exists in a loop and are dependent on loop-invariant
values are also recorded. Here, the loop itself can be aggressively unrolled and
the \ALLS\ can be merged or replaced with one of the methods from the custom 
allocator. It is also important that the size is dependent on a loop-invariant 
value, or else the compiler cannot affirmatively analyze multiples of the size or
determine an appropriate unroll factor (unrolling too much without knowledge of
the \SIZE\ may be sub-optimal). An example of this is shown in Line 8 of Figure ~\ref{fig:sizes}

@@

\subsection{Custom Allocator} \label{ca}
The final component of the system is a custom memory allocator, managed by both
the compiler and runtime and designed with consideration of FaaS characteristics.
In particular, the allocator aims to reduce raw \ALL\ and \DALL\ time, consolidate 
\ALLS\ , and utilize compiler analysis to optimize dynamic memory overheads.

The allocator primarly takes the static analyses (see \ref{sa})
as input to replace as many \ALLS\ and \DALLS\ with invocations to the allocator. 
First, \ALLS\ with statically analyzable \SIZES\ are prime candidates to be 
\textit{pre-allocated} because all the information needed to allocate is known at
compile-time. These \ALLS\ are placed into a compiler-directed memory pool that is
allocated by the runtime in parallel to FaaS instance startup and before the entry-point function 
of the workload, $W$, begins. The size of the pool is the sum of all statically 
analyzable \SIZES\ of the selected \ALLS . As a result, the \ALL\ function calls are 
replaced by load of a compiler-determined offset of the base address of the pool.
An \ALL\ can only be selected for this transformation if it is not flagged by 
the escapes analysis (see the third analysis in \ref{sa}) \textit{and} the compiler 
can deduce the number of invocations of its parent function. Otherwise, the data
in memory or the pointer itself may be overwritten accidentally. The latter condition
can be alleviated with aggressive inlining and function versioning. An example of 
this transformation is shown in Figure~\ref{fig:pool}. There are a number of 
benefits to this approach: 1) alleviating function call overheads, 2) enabling 
prefetching of allocation pointer address, 3) hiding allocation overhead by performing 
it before $W$ even begins.

\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.77\textwidth]{figs/pool.png} 
        \caption{Example allocations (top) transformed to fetch from the compiler-directed
        memory pool (bottom; here, wrapped into the function \texttt{pool\_fetch}). The 
        basic functionality of the pool is shown to the right. In context of simplified LLVM 
        IR and \textit{libc} \texttt{malloc}}  
	    \label{fig:pool}
    \end{minipage}\hfill
\end{figure}
